---
title: "7 CFA: Second Order"
author: "Rushabh BArbhaya"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Hierarchical CFA Practice Assignment

## DASS - Depression, Anxiety, and Stress Scale

The DASS is a measurement scale that examines the depression, anxiety, and stress of an individual. Here are the DASS-21 items:

    1 I found it hard to wind down.
    2 I was aware of dryness of my mouth.
    3 I couldn't seem to experience any positive feeling at all.
    4 I experienced breathing difficulty.
    5 I found it difficult to work up the initiative to do things.
    6 I tended to over-react to situations.
    7 I experienced trembling (eg, in the hands).
    8 I felt that I was using a lot of nervous energy.
    9 I was worried about situations in which I might panic and make a fool of myself.
    10 I felt that I had nothing to look forward to.
    11 I found myself getting agitated.
    12 I found it difficult to relax.
    13 I felt down-hearted and blue.
    14 I was intolerant of anything that kept me from getting on with what I was doing.
    15 I felt I was close to panic.
    16 I was unable to become enthusiastic about anything.
    17 I felt I wasn't worth much as a person.
    18 I felt that I was rather touchy.
    19 I was aware of the action of my heart in the absence of physical exertion.
    20 I felt scared without any good reason.
    21 I felt that life was meaningless.

The scale should be scored using the following latent variables:

    Depression: Questions 3, 5, 10, 13, 16, 17, 21
    Anxiety: Questions 2, 4, 7, 9, 15, 19, 20
    Stress: 1, 6, 8, 11, 12, 14, 18

## Load the Data and Libraries

In this section, load the assignment data and the libraries you need for this assignment. 

```{r}
library(lavaan)
library(semPlot)

data = read.csv("assignment_secondcfa.csv")
```

## Build Models

Create a three-factor model of the DASS as your measurement model.

```{r}
model1 = '
depression =~ Q3 + Q5 + Q10 + Q13 + Q16 + Q17 + Q21
anxiety =~ Q2 + Q4 + Q7 + Q9 + Q15 + Q19 + Q20
stress =~ Q1 + Q6 + Q8 + Q11 + Q12 + Q14 + Q18 
'
```

Create a second-order model of the DASS with one general factor that predicts depression, anxiety, and stress. 

```{r}
model2 = '
depression =~ Q3 + Q5 + Q10 + Q13 + Q16 + Q17 + Q21
anxiety =~ Q2 + Q4 + Q7 + Q9 + Q15 + Q19 + Q20
stress =~ Q1 + Q6 + Q8 + Q11 + Q12 + Q14 + Q18
global =~ depression + anxiety + stress
'
```

Create a bi-factor model of the DASS with one general factor that predicts the items on the DASS. 

```{r}
model.bi = '
depression =~ Q3 + Q5 + Q10 + Q13 + Q16 + Q17 + Q21
anxiety =~ Q2 + Q4 + Q7 + Q9 + Q15 + Q19 + Q20
stress =~ Q1 + Q6 + Q8 + Q11 + Q12 + Q14 + Q18 
global =~ Q3 + Q5 + Q10 + Q13 + Q16 + Q17 + Q21 + Q2 + Q4 + Q7 + Q9 + Q15 + Q19 + Q20 + Q1 + Q6 + Q8 + Q11 + Q12 + Q14 + Q18
'
```

## Analyze Models

Analyze the measurement model of the DASS. 

```{r}
model1.fit = cfa(model1, data = data)
```

Analyze the second-order model of the DASS. 

```{r}
model2.fit = cfa(model2, data = data)
```

Analyze the bi-factor model of the DASS. You will need to add `check.gradient = F` to the cfa function to have this model converge correctly. 

```{r}
model.bi.fit = cfa(model.bi, data = data, check.gradient = F, std.lv = T, orthogonal = T)
```

Oh no! Your model has a negative variance. This result means this model hasn't run correctly. We will deal with these more later, so ignore it for now. 

## Summarize Models

Summarize the measurement model of the DASS. 

```{r}
summary(model1.fit, standardized = T, rsquare = T, fit.measures = T)
fitMeasures(model1.fit)
```

Summarize the second-order model of the DASS. 

```{r}
summary(model2.fit, standardized = T, rsquare = T, fit.measures = T)
fitMeasures(model2.fit)
```

Summarize the bi-factor model of the DASS. 

```{r}
summary(model.bi.fit, standardized = T, rsquare = T, fit.measures = T)
fitMeasures(model.bi.fit)
```

QUESTION: Does the measurement model (the first one) appear to be a good model? Use both the loadings and fit indices to answer this Question. 
ANSWER: All the variances are positive; standard loadings are greater than $\pm$ 0.3; Error values are don't look to high. SRMS is also quite low; All other fit indices also look decently good. Overall it's a good model.

QUESTION: Are the correlations high in the measurement model to support a second order model? 
ANSWER: Yes

QUESTION: Examine the summaries of each model. Does a second order or bi-factor model appear to fit the DASS better? 
ANSWER: Second order model is a better model as compared to bi-factor model. There are a lot of loadings which have values < $\pm$ 0.3

QUESTION: Do the Questions load on to the domain specific factors in the bi-factor model? 
ANSWER: No

QUESTION: Do the Questions load on to the general factor in the bi-factor model?
ANSWER: No