---
title: "Assignment 1 - 1 Intro R"
author: "Rushabh Barbhaya"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Create R chunks to display both the output and code in this file. For each of the following sections, create the code to answer the question and output the answer (if necessary).

#	Make a vector of numbers. 
    o	Use the rep or seq function.
    o	Include a concatenate c().

```{r vector}
vector1 = rep(0, 10)
cat("repeat --> ", vector1)
vector2 = seq(1,10,1)
cat("sequence --> ", vector2)
vector3 = c(vector1, vector2)
cat("concat --> ", vector3)
```


#	Use the length() function to figure out how long your vector is.
```{r len}
cat("Length of the vector1 variable --> ", length(vector1))
cat("Length of the vector2 variable --> ", length(vector2))
cat("Length of the vector3 variable --> ", length(vector3))
```


#	Make a matrix at least 5 by 5. 
    o	Pull out the value for row 2 column 3.
```{r Matrix}
mat1 = matrix(0, 5, 5)
diag(mat1) = 1
mat1

cat("value for row 2 column 3 --> ", mat1[2,3])
```


#	Use the head() function to look at the dataset USArrests. 
    o	Calculate a summary of USArrests.
```{r dataset}
USArrests <- read.csv("1 intro R.csv")
head(USArrests)
```

#	Import a data set (attached). 
    o	Recode all the values to LEFT and RIGHT in the whichhand column.
    o	Create a summary of the dataset.
    o	Exclude all navalues in the dataset.
    o	Create a dataset with only LEFT handed participants.
    o	Calculate the correlation table of just the following variables: LR_switch, finger_switch, rha, word_length, letter_freq

```{r recode}
# levels(USArrests$whichhand)
# rep("Right", 7)
USArrests$whichhand = factor(USArrests$whichhand,
                   levels = c("", "left", "Left", "right", "Right", "RIGHT", "right ", "Right ",
                        "right hand", "Right.", "right_left", "with Cap letter"),
                   labels = c(NA, "Left", "Left", "Right", "Right", "Right", "Right", "Right", 
                              "Right", "Right", "Both", NA))
USArrests$whichhand = droplevels(USArrests$whichhand)
# Thank you for this technique Professor
summary(USArrests)

cat("Dimensions With NA Values: ", dim(USArrests))
USArrests = na.omit(USArrests)
cat("Dimensions Without NA Values: ", dim(USArrests))

leftOnly = USArrests[USArrests$whichhand == "Left",]
summary(leftOnly)

cor(USArrests[8:12])
#cor(USArrests[8:12], use = "pairwise.complete.obs") This gave the same result as above
```


#	Bonus: 
    o	Run a regression using the following: lm(rating~finger_switch, data=data) **or whatever you called the data
    o	Save the output of the regression.
    o	Pull out only the slope from the coefficients.
    
```{r regression}
reg = lm(rating ~ finger_switch, data = USArrests)
reg$coefficients
reg$coefficients[1]
reg$coefficients[2]
```

