---
title: "11 MGCFA"
author: "Rushabh Barbhaya"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## MGCFA Class Assignment

Using the same RS14 dataset from class, examine the multigroup analysis using `Ethnicity` as the grouping variable of interest. 

## Data Libraries

Load the included dataset and libraries for this assignment. 

```{r}
library(lavaan)
library(rio)
library(semPlot)

data = import("assignment_mgcfa.csv")
table(data$Ethnicity)
```

## Data Cleanup

In this example, code 1 as Black and 2 as White. Be sure to first change the variable to a factor with labels. 

```{r}
data$Ethnicity = factor(data$Ethnicity, levels = c(1,2), labels = c("Black", "White"))
data = subset(data, !is.na(Ethnicity))
```

## Overall CFA

In this section, build the overall one-factor CFA for the RS14. You can use the example we did in class to get started. Include the `cfa()` and `summary()` of your model here. You should call this model `overall.model`.

```{r}
overall.model = '
RS =~ RS1 + RS2 + RS3 + RS4 + RS5 + RS6 + RS7 + RS8 + RS9 + RS10 + RS11 + RS12 + RS13 + RS14
'
overall.fit = cfa(model = overall.model, data = data, meanstructure = T)
summary(overall.fit, standardized = T, rsquare = T, fit.measure = T)
semPaths(overall.fit, whatLabels = "std", layout = "spring", edge.label.cex = 1)
```

## White Overall Summary

Create a `white.fit` of the model for only White participants. Include the `summary()` of this model. 

```{r}
white.fit = cfa(model = overall.model, data = data[data$Ethnicity == "White",], meanstructure = T)
summary(white.fit, standardized = T, rsquare = T, fit.measure = T)
semPaths(white.fit, whatLabels = "std", layout = "spring", edge.label.cex = 1)
```

## Black Overall Summary

Create a `black.fit` of the model for only Black participants. Include the `summary()` of this model. 

```{r}
black.fit = cfa(model = overall.model, data = data[data$Ethnicity == "Black",], meanstructure = T)
summary(black.fit, standardized = T, rsquare = T, fit.measure = T)
semPaths(black.fit, whatLabels = "std", layout = "spring", edge.label.cex = 1)
```

## Add the fit to table

In the code below, we've started a table for our fit indices. Fill in the section labeled `#update here` with the same `round(fitmeasures(...))` code used in the first row. 

```{r}
library(knitr)
table_fit <- matrix(NA, nrow = 8, ncol = 6)
colnames(table_fit) = c("Model", "X2", "df", "CFI", "RMSEA", "SRMR")
table_fit[1, ] <- c("Overall Model", round(fitmeasures(overall.fit, 
                                                 c("chisq", "df", "cfi",
                                                   "rmsea", "srmr")),3))
kable(table_fit)

table_fit[2, ] <- c("White Model", round(fitmeasures(white.fit, 
                                                 c("chisq", "df", "cfi",
                                                   "rmsea", "srmr")),3))

table_fit[3, ] <- c("Black Model", round(fitmeasures(black.fit, 
                                                 c("chisq", "df", "cfi",
                                                   "rmsea", "srmr")),3))

kable(table_fit)
```

QUESTION: Does your model appear to fit ok overall and within each individual group? \
ANSWER: Both the groups fit well. Black Ethnicity Model seems to fit better as compared to White Ethnicity Model

## Configural Invariance

Create a `configural.fit` model using `Ethnicity` as your group variable. Include a `summary()` of this model. Add in the `configural.fit` to your table. 

```{r}
configural.fit = cfa(model = overall.model, data = data, meanstructure = T, group = "Ethnicity")
summary(configural.fit, standardized = T, rsquare = T, fit.measure = T)

table_fit[4, ] <- c("Configural Model", round(fitmeasures(configural.fit, 
                                                 c("chisq", "df", "cfi",
                                                   "rmsea", "srmr")),3))

kable(table_fit)
```

## Metric Invariance

In this section, create `metric.fit`, summarize the model, and add the model information to your table. 

```{r}
metric.fit = cfa(model = overall.model, data = data, meanstructure = T, group = "Ethnicity", group.equal = c("loadings"))
summary(metric.fit, standardized = T, rsquare = T, fit.measure = T)

table_fit[5, ] <- c("Metric Model", round(fitmeasures(metric.fit, 
                                                 c("chisq", "df", "cfi",
                                                   "rmsea", "srmr")),3))

kable(table_fit)
```

## Scalar Invariance

In this section, create `scalar.fit`, summarize the model, and add the model information to your table.  

```{r}
scalar.fit = cfa(model = overall.model, data = data, meanstructure = T, group = "Ethnicity", group.equal = c("loadings", "intercepts"))
summary(scalar.fit, standardized = T, rsquare = T, fit.measure = T)

table_fit[6, ] <- c("Scalar Model", round(fitmeasures(scalar.fit, 
                                                 c("chisq", "df", "cfi",
                                                   "rmsea", "srmr")),3))

kable(table_fit)
```

## Strict (Error) Invariance

Last, add `strict.fit` by including error variances in your multigroup model, summarize that model, and include the fit information in the table. 

```{r}
strict.fit = cfa(model = overall.model, data = data, meanstructure = T, group = "Ethnicity", group.equal = c("loadings", "intercepts", "residuals"))
summary(strict.fit, standardized = T, rsquare = T, fit.measure = T)

table_fit[7, ] <- c("Strict Model", round(fitmeasures(strict.fit, 
                                                 c("chisq", "df", "cfi",
                                                   "rmsea", "srmr")),3))

kable(table_fit)
```

QUESTION: Looking at your results, do you see anywhere that indicates non-invariance (i.e., has a larger change than CFI = .01)? Where did it appear to break down? \
ANSWER: Yes, it broke when building Scalar Model

## Partial Invariance

Use the code provided from class to calculate the potential pathway that should be freed to improve our model fit for partial invariance. 

```{r}
partial_syntax = paste(colnames(data)[4:17], "~~", colnames(data)[4:17])
CFI_list  <- 1:length(partial_syntax)
names(CFI_list) <- partial_syntax

for (i in 1:length(partial_syntax)){
  temp <- cfa(model = overall.model, 
              data = data,
              meanstructure = TRUE,
              group = "Ethnicity",
              group.equal = c("loadings", "intercepts", "residuals"),
              group.partial = partial_syntax[i])
  CFI_list[i] <- fitmeasures(temp, "cfi")
}
which.max(CFI_list)
```

## Update the Model

Create `partial.fit` by updating your model to allow for the freed parameter. Summarize the model, and add this model's fit information to your table. 

```{r}
partial.fit <- cfa(model = overall.model, data = data, meanstructure = TRUE, group = "Ethnicity", group.equal = c("loadings", "intercepts", "residuals"), group.partial = c("RS3 ~~ RS3"))

summary(partial.fit, standardized = T, rsquare = T, fit.measure = T)

table_fit[8, ] <- c("Partial Model", round(fitmeasures(partial.fit, c("chisq", "df", "cfi", "rmsea", "srmr")),3))

kable(table_fit)
```

QUESTION: Does the freeing of this one parameter bring your model up to partial invariance for that step? \
ANSWER: Yes, it brought it close.

QUESTION: What does the result of the partial invariance step imply for interpreting this scale? \
ANSWER: It indicates that the model is different from the one before. And significantly at that.


