---
title: "6 CFA: Basics"
author: "YOUR NAME"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Confirmatory Factor Analysis Practice Assignment

In the section below, you should load the `lavaan` and `semPlot` libraries. The data is included in `assignment_cfabasics.csv`, and you should load that dataset. 

```{r warning=FALSE, message=FALSE}
library(lavaan)
library(semPlot)
library(broom)

data <- read.csv("assignment_cfabasics.csv")
```

## The Data

The data represents a 30-item questionnaire called CAAFI: Computer Aversion, Attitudes, and Familiarity Index. This questionnaire includes items on computer usage, attitudes, and familiarity. 

- Computer Familiarity: Items 3, 13-14, 16, 20-23, 27, and 30. 
- Computer Attitudes: Items 1-2, 4-5, 8, 11, 18-19, and 28-29.
- Computer Aversion: Items 6-7, 9-10, 12, 15, 17, and 24-26.

Here are the items that are present on the survey: 

1. I enjoy using computers.
2. Being able to use a computer is important to me.
3. I keep up with the latest computer hardware.
4. Computers are beneficial because they save people time.
5. I like using word-processing programs.
6. I feel like a fool when I am using a computer and others are around.
7. I am smart enough to use a computer.
8. I avoid using computers whenever possible.
9. I do not understand how to use computer software (e.g., word-processing
programs, spreadsheet programs, etc.).
10. I feel that I understand how to use computer files, documents, and folders.
11. I use a computer input device every day (e.g., a keyboard, a touch pad, a mouse).
12. I can use a computer to successfully perform tasks.
13. I can add new hardware to a computer.
14. I enjoy reading computer magazines.
15. When I use a computer, I am afraid that I will damage it.
16. I enjoy connecting new computer accessories.
17. I must have a reference manual or a help file to run computer software.
18. E-mail is an easy way to communicate with people.
19. I use e-mail every day.
20. I am comfortable changing (installing/upgrading) computer software.
21. I often read computer books.
22. My friends often ask me computer-related questions.
23. I often read computer magazines.
24. Overall, I feel that I don’t know how to use a computer.
25. Computers are too scientific for me.
26. When using a computer, I often lose data.
27. I enjoy learning to use new software programs.
28. I like to use computer input devices such as a keyboard, a touch pad, a mouse,
etc.
29. Using a computer is entertaining.
30. I keep up with the latest computer software.

## Build a Model

In this section, create a one-factor model of all of the items together. 

```{r}
# for (i in seq(30)){
#   cat(paste0('q',i,'+'))
# }

model.one = '
computer =~ q1+q2+q3+q4+q5+q6+q7+q8+q9+q10+q11+q12+q13+q14+q15+q16+q17+q18+q19+q20+q21+q22+q23+q24+q25+q26+q27+q28+q29+q30
'
```

In this section, create a three-factor model based on the items proposed grouping listed above.

```{r}
# - Computer Familiarity: Items 3, 13-14, 16, 20-23, 27, and 30. 
# - Computer Attitudes: Items 1-2, 4-5, 8, 11, 18-19, and 28-29.
# - Computer Aversion: Items 6-7, 9-10, 12, 15, 17, and 24-26.

model.three = '
familiarity =~ q3 + q13 + q14 + q16 + q20 + q21 + q22 + q23 + q27 + q30
attitudes =~ q1 + q2 + q4 + q5 + q8 + q11 + q18 + q19 + q28 + q29
aversion =~ q6 + q7 + q9 + q10 + q12 + q15 + q17 + q24 + q25 + q26
'
```

## Analyze Your Model

Analyze both the one-factor and three-factor model.

```{r}
model.one.fit = cfa(model.one, data = data)
model.three.fit = cfa(model.three, data = data)
```

## Summarize Your Model

Include a summary of the one-factor and three-factor model with standardized output, rsquare values, and fit.measures. 

```{r}
summary(model.one.fit, standardize = T, fit.measures = T, rsquare = T)
summary(model.three.fit, standardize = T, fit.measures = T, rsquare = T)
```

## Diagram Your Model

Include a diagram for the three-factor model. Use the `what` and `edge.color` arguments to denote the strongest paths. 

```{r}
semPaths(model.one.fit, whatLabels = 'std', what = 'std', layout = 'tree', edge.color = 'blue', edge.label.cex = 1)
semPaths(model.three.fit, whatLabels = 'std', what = 'std', layout = 'tree', edge.color = 'orange', edge.label.cex = 1)
```

## Compare Models 

Use the `anova()` function and the `fitmeasures()` to create a model comparison. 

```{r}
anova(model.one.fit, model.three.fit)
fitmeasures(model.one.fit, c('aic', 'ecvi'))
fitmeasures(model.three.fit, c('aic','ecvi'))
```

QUESTION: Which model is better? 
ANSWER: model.three has a lower chisq value, aic and bic score as compared to model.one $\therefore$ model.three is better. model.three is also significantly different from model.one

## Clean Output

Use either `model_parameters` or `broom`'s `tidy` function to neatly print out the model parameters for the three-factor model.

```{r}
tidy(model.one.fit)
tidy(model.three.fit)
```


QUESTION: Looking at the item loadings, are any of the items poor (i.e., they do not strongly relate to their factor)?
ANSWER: Almost all the loadings look good. There are some which are strongly related. q14 and q21 in model with one variable. 

```{r}
correl = residuals(model.three.fit, type = 'cor')
```


QUESTION: Looking at the correlation table, do you think these are three separate factors? 
ANSWER: There are some questions which are greately correlated together.

## Modifications

Include the potential modification indices for the three-factor model. Be sure to use the `sort` option to see the strongest modification indices. 

```{r}
modindices(model.one.fit, sort. = T)
```

QUESTION: What potential indices do you see in these modification indices? What questions may be asking the same material twice? 
ANSWER: q14 and q23 being correlated. If that correlation is removed the chisq value can be moved by 428 points. Similar cases for others as well.